### 2020-10-20 authors.tex
updated authors list

\author{
Joshua~L.~Steffen\altaffilmark{1}, Hai~Fu\altaffilmark{1,2}, J.~M.~Comerford\altaffilmark{3}, Y.~Sophia~Dai\altaffilmark{4}, Shuai~Feng\altaffilmark{5}, Cheng~Li\altaffilmark{6}, and Rui~Xue\altaffilmark{1}
}
\altaffiltext{1}{Department of Physics \& Astronomy, University of Iowa, Iowa City, IA 52242}
\altaffiltext{2}{Insititute for Astronomy, University of Hawaii, Honolulu, HI 96822}
\altaffiltext{3}{Department of Astrophysical and Planetary Sciences, University of Colorado, Boulder, CO 80309}
\altaffiltext{4}{Chinese Academy of Sciences South America Center for Astronomy (CASSACA), 20A Datun Road, Beijing 100012, China}
\altaffiltext{5}{Key Laboratory for Research in Galaxies and Cosmology, Shanghai Astronomical University, Chinese Academy of Sciences}
\altaffiltext{6}{Tsinghua Center for Astrophysics and Physics Department, Tsinghua University, Beijing 100084, China}