### 2021-01-20 table1.tex

% ApJ Format
%\documentclass[apj,twocolumn]{emulateapj}
%\usepackage{apjfonts}

%\begin{document}

\begin{table}
\begin{center}
\caption{Close Galaxy Pairs and Multiples in MaNGA IFUs}
\label{tab:coords}
\begin{tabular}{lc ccc}
\hline
\hline
Plate-IFU & Index & R.A. (J2000) & Decl. (J2000) & Redshift \\
          &           & (deg)        & (deg)         &     \\
\hline
\input{table1_data.tex}
\hline
\end{tabular}
\end{center}
\tablecomments{
This table lists a total of 404 plate-IFUs that contain close galaxy pairs and multiples in the eighth MaNGA Product Launch (MPL-8). The second column lists the component index, where ``0'' indicates the primary target of the MaNGA observation. For each IFU, all components within $\pm$2000\,km~s$^{-1}$ of the primary are listed, sorted in ascending angular distance from the primary.\\
(This table is available in its entirety in a machine-readable form in the online journal. A portion is shown here for guidance regarding its form and content.)
}
\end{table}

%\end{document}